---
title: "WB Survey Data"
description: |
  The White Box Survey data was collected from participating examiners by administering a post-study survey to gather personal information (i.e. sex, age) and occupational information (i.e. experience, certifications). The survey also asked examiners about how difficult they found this study to complete overall. Unfortunately, the survey data does not have any examiner IDs attached to the responses so it could not be used in conjunction to the Determinations data set.
output: 
  distill::distill_article:
    code_folding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(tidyverse)
library(readxl)
library(janitor)
library(openintro)
library(ggridges)
library(gcookbook)
library(gridExtra)
library(dplyr)
library(ggthemes)
wbstudy <- read_excel("WhiteBox_PublicDataRelease.xlsx", sheet = 1)
```

According to [Appendix SI-7][WBpaper] of the study, 170 examiners participated in the study while 169 responded to the background survey. 82% of these 170 participants are from the United States. I was interested in how the sex and education levels of these 169 participants impact how difficult they found this study. I am also curious about how representative this sample is of the population of fingerprint examiners. The variable names that I will be referencing to are in the table below.

```{r cleaning data, include=TRUE}
wb_clean = wbstudy %>%
  clean_names()

head(wb_clean)
```
**Sex Distribution of WB Participants**

First, I wanted to know the sex distribution of the respondents. To my surprise, there were more female respondents than there are male respondents. According to [Zippia][zip]'s data on fingerprint examiner demographics, about 56% of examiners are female and 3.1% of examiners hold a masters degree. [CSI Accelerator][CSI], an education resource website for forensics students, states that the most common path to becoming a fingerprint examiner is by earning a bachelor's degree. Additionally, in 2019 NPR wrote that the majority of college-educated workers will soon be women.

```{r gender dist, include=TRUE}
wb_clean$x1_sex<- factor(wb_clean$x1_sex,levels = c("Female", "Male"))

sex <- ggplot(wb_clean, aes(x1_sex, fill = x1_sex)) +
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "white") +
  labs(title = "WB Survey Sex Distribution", x = "Sex", y = "Count") +
  scale_fill_manual("Sex", values = c("lightblue2", "steelblue")) +
  theme_minimal()

sex
```
Thus, it is not too surprising that that the majority of the participants in this study are female. It actually means that this sample is probably representative of what the sex distribution looks like. 

**Education Level of Participants**

```{r education, include = TRUE}
wb_clean$x3_education<- factor(wb_clean$x3_education,levels = c("HS", "Assoc", "Bachelor", "Grad"))

edu_sex <- ggplot(wb_clean, aes(x = x1_sex, fill = x3_education)) +
  geom_bar(position = "fill") +
  labs(title = "WB Survey Education Level by Sex", x = "Degree", y = "Count") +
  scale_fill_manual("Sex", values = c("lightblue1", "lightskyblue", "steelblue2", "dodgerblue4")) +
  theme_minimal()

edu_sex

```
It was surprising to me to see a female-majority in a field that I expected to be male-dominant. This opened the door to many questions since I have never been in a majority-female academic space or field yet. I know from past experience that male-dominant fields (especially in the sciences) can get very discouraging and frustrating and for me, I ended up turning away from my initial major because of such culture. My next question was how education level varies between genders and if it has anything to do with how difficult the participants found the exercises. I did not know what to expect, since I was used to seeing and hearing about men being more highly educated in male-majority fields.

**Computer Proficiency & Experience**

```{r computer time}
wb_clean$x18_compare_to_casework<- factor(wb_clean$x18_compare_to_casework, levels = c("	
MUCH EASIER than casework", "EASIER than casework", "SIMILAR to casework", "HARDER than casework", "MUCH HARDER than casework"))

wb_clean$x11_computer_based<- factor(wb_clean$x11_computer_based, levels = c("0% computer", "1-30% computer", "30-60% computer", "60-99% computer", "100% computer"))

comp_time_diff <- ggplot(wb_clean, aes(x = x11_computer_based, fill = x18_compare_to_casework)) +
  geom_bar() +
  labs(title = "WB Survey Difficulty Rating by Computer Time", x = "Computer Time Spent", y = "Count") +
  scale_fill_manual("Difficulty", values = c("lightblue1", "lightskyblue", "steelblue2", "dodgerblue4", "midnightblue")) +
  theme_minimal()

comp_time_diff + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

[WBpaper]: file:///Users/azhang/Downloads/pone.0110179.s007.pdf "White Box Paper"
[zip]: https://www.zippia.com/latent-fingerprint-examiner-jobs/demographics/ "Fingerprint examiner job demographics"
[CSI]: https://csiaccelerator.org/careers/fingerprint-analyst/ "CSI Accelerator's how to become a fingerprint examiner webpage"
